Random Forests", BUAs

### 2024-02-12 src/bua_RF.Rmd
bua

df <- df %>% 
  group_by(year) %>% 
  mutate(n.slag = lag.listw(knn.l, n)) %>% 
  group_by(BUA22CD) %>% 
  mutate(n.l.slag = dplyr::lag(n.slag, n=1, order_by=year),
         n.London.lag = dplyr::lag(n.London, n=1, order_by = year),
         n.nearest.city.lag = dplyr::lag(n.nearest.city, n=1, order_by = year))
```

@@

NOT DONE

@@

# Lookup: OA to regions
path.lookup <- paste0(path, "/data/raw/Output_Area_to_Region_(December_2011)_Lookup_in_England.csv")  
lookup.region <- read_csv(path.lookup) %>% dplyr::select(OA11CD, RGN11CD, RGN11NM)

df <- df %>% left_join(lookup.region, by = c("oa11cd" = "OA11CD")) %>% 
  relocate(c(RGN11NM, RGN11CD), .after=year) %>% 
  mutate(help = substring(oa11cd, 1, 1),
         RGN11CD = ifelse(help=="E", RGN11CD,
                         ifelse(help=="W", "Wales",
                                ifelse(help == "S", "Scotland", "NI"))),
         RGN11NM = ifelse(help=="E", RGN11NM,
                         ifelse(help=="W", "Wales",
                                ifelse(help == "S", "Scotland", "NI")))) %>% 
  dplyr::select(-help)

@@

# For growth:
df <- df %>% group_by(oa11cd) %>% 
  mutate(#n.lag = dplyr::lag(n),
         growth = log(n/n.lag),
                    abs.growth = n - n.lag) %>% 
  relocate(n.lag, .after = n)

# out.path <- paste0(path, "/data/temp/df_for_oa_rf.csv")
# df <- read_csv(out.path)
# df %>% write_csv(file = out.path)

@@

# remove not used columns
df <- df %>% dplyr::select(-c("Birmingham", "Glasgow", "Liverpool", "Bristol",
                        "Manchester", "Sheffield", "Leeds", "Edinburgh",
                        "Leicester", "dist.city.name", "dist.retail.name", "name"))

# There are no lags for 1996, so it is removed
df <- df %>% filter(year > 1996) %>% 
  dplyr::select(BUA22CD, BUA22NM, year, n, n.l.slag, n.nearest.city.lag, 
                n.London.lag, City.of.Westminster, dist,) %>% 
  rename(London = City.of.Westminster)

df %>% group_by(year) %>% 
  filter(n == 0) %>% 
  count(n)
# in 2012, 120848 0s

@@

```{r}
set.seed(123)
indices <- CreateSpacetimeFolds(df, 
                                spacevar = "BUA22CD",
                                timevar = "year", 
                                k = 10)

@@

start.time <- Sys.time()
model.all <- train(n ~ #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + n.nearest.city + year,
                     n.London.lag + n.nearest.city.lag + n.l.slag + year + London + dist ,
                     #London.v + city.v + retail.v + n.l.slag + year,
                   data = df, 
                   trControl = tc,
                   method = "ranger",
                   na.action = na.omit,
                   #preProc = c("center", "scale"),
                   importance = "impurity")

# stopCluster(cl)
end.time <- Sys.time()
time.taken <- round(end.time - start.time,2)
time.taken #7.3m

@@

# Random Forest 
# 
# 134528 samples
#      6 predictor
# 
# No pre-processing
# Resampling: Cross-Validated (10 fold) 
# Summary of sample sizes: 113505, 105938, 105938, 113520, 113490, 105938, ... 
# Resampling results across tuning parameters:
# 
#   mtry  splitrule   RMSE      Rsquared   MAE     
#   2     variance    79.06941  0.6270099  24.31033
#   2     extratrees  93.85626  0.5098337  26.64545
#   4     variance    75.36044  0.6507424  21.62872
#   4     extratrees  85.91194  0.5748039  25.88575
#   6     variance    75.32465  0.6476521  21.19409
#   6     extratrees  83.85574  0.5904021  26.05747
# 
# Tuning parameter 'min.node.size' was held constant at a value of 5
# RMSE was used to select the optimal model using the smallest value.
# The final values used for the model were mtry = 6, splitrule = variance and min.node.size = 5.

varimp_mars <- varImp(model.all) 
varimp_mars$importance <- varimp_mars$importance %>% 
  rownames_to_column() %>% 
  mutate(rowname = ifelse(rowname == "dist", "distance to the nearest city",
                          ifelse(rowname == "dist.retail", "distance to the nearest retail centre", 
                                 ifelse(rowname == "London", "distance to London",
                                        ifelse(rowname == "n.nearest.retail.lag", "Nearest retail centre's wensite density, t-1",
                                               ifelse(rowname == "n.London.lag", "London's wensite density, t-1",
                                                      ifelse(rowname == "n.l.slag", "spatial and temporal lag of wensite density",
                                                             ifelse(rowname == "n.nearest.city", "Nearest city's wensite density, t-1", rowname)))))))) %>% 
  column_to_rownames()

path.out <- paste0(path, "/outputs/rf/figures/varimp_OA.png")
ggplot(varimp_mars) + theme_minimal() + labs(x="") #+ theme(panel.background = element_rect(fill = "white"))
ggsave(path.out)

path.out <- paste0(path, "/outputs/rf/models/oa_rf/all_OAs/all_OAs.RData")
save(model.all, file = path.out)

# ffsmodel_LLO <- ffs(predictors = df[, c("area", "dist", "dist.retail", "London", "south",  
#                                       "n.slag", "n.lag", "n.l.slag", "n.nearest.city", 
#                                                         "year")], 
#                     response = df$n, #%>% dplyr::select(n) %>% as.data.frame %>% mutate(n = as.numeric(n)),
#                     trControl = tc,
#                     method = "ranger",
#                     #na.action = na.omit,
#                     #importance = "impurity",
#                     metric="Rsquared",
#                     verbose=FALSE)
# ffsmodel_LLO
# ffsmodel_LLO$selectedvars
# plot_ffs(ffsmodel_LLO)
```

@@

@@

# detectCores()
# cl <- makePSOCKcluster(2)
# registerDoParallel(cl)

@@

regions <- df %>% ungroup() %>% distinct(RGN11CD) 
regions <- as.vector(regions$RGN11CD)

@@

  set.seed(71)
  indices <- CreateSpacetimeFolds(df %>% filter(RGN11CD!=i), 
                                spacevar = "oa11cd",
                                timevar = "year", 
                                k = k)

@@

  model.all <- train(n ~ n.London.lag + n.nearest.city.lag + n.nearest.retail.lag + n.l.slag + year + London + dist + dist.retail,
                   data = df %>% filter(RGN11CD!=i), 
                   trControl = tc,
                   method = "ranger",
                   na.action = na.omit,
                   #preProc = c("center", "scale"),
                   importance = "impurity")
  #assign(paste0("model.all",as.character(i)), model.all)

  #file <- paste0(path, "/outputs/rf/models/oa_rf/regions/region_", as.character(i), ".RData")
  file <- paste0("/hdd/tmp/regions/region_", as.character(i), ".RData")  
  save(model.all, file = file)  # file name = region not in the training data
  rm(model.all)

@@

# clear up
rm(model.all) # this is the last item from the loop
rm(list = ls(pattern = "^region_"))

@@

df.predictions <- data.frame()
for (i in regions){
  #for (j in regions[!regions == i]){
    pred <- predict(fit.model.all[names(fit.model.all)==paste0("region_", i)], 
                    df[df$RGN11CD==i,])
    pred <- as.data.frame(pred)
    rownames(pred) <- c()
    assign(paste0("region.predict.model.", ".on.", i), pred)

    d <- paste0("region.predict.model.", ".on.", i)
    d <- cbind(get(d), df %>% filter(RGN11CD==i)) #%>% arrange(year, ladcd)

    #***dplyr::select(RGN21CD, ladcd, year, growth, 1)***

    d <- d[,c(5, 2, 4, 6, 1)]     
    colnames(d)[5] <- "predictions"
    d <- d %>% mutate(trained.on = i)
    df.predictions <- rbind(d, df.predictions)
  #}
}

df.predictions <- df.predictions %>% mutate(test.train = paste0(RGN11CD, ".", trained.on))

@@

path.lookup <- paste0(path, "/data/raw/Local_Authority_District_to_Region_(April_2021)_Lookup_in_England.csv")  
lookup.region <- read_csv(path.lookup) %>% 
  dplyr::select(RGN21CD, RGN21NM) %>% 
  distinct() %>% 
  rename(RGN11CD = RGN21CD) %>% 
  add_row(RGN11CD = c("Scotland", "N0", "Wales"), 
          RGN21NM = c("Scotland", "Nortern Ireland", "Wales"))

path.out <- paste0(path, "/outputs/rf/heatmap_OA.png")
rf.year.all.metrics %>% 
  as.data.frame() %>%
  rownames_to_column(var = "metrics") %>% 
  pivot_longer(!metrics, names_to = "train.test") %>% 
  pivot_wider(names_from = metrics, values_from = value) %>% 
  separate(train.test, c("train", "test"), remove = F) %>% 
  left_join(lookup.region, by = c("train" = "RGN11CD")) %>% 
  rename(train.region=RGN21NM) %>% 
  left_join(lookup.region, by = c("test" = "RGN11CD")) %>% 
  rename(test.region=RGN21NM) %>% 
  arrange(Rsquared) %>% 
  dplyr::select(train.region, test.region, Rsquared) %>% 
  ggplot(aes(train.region, test.region, fill= Rsquared)) + 
  geom_tile() +
  coord_fixed() +
  labs(y = "", x = "") + #y: test.region, x: train.region
  scale_fill_gradientn(colours = terrain.colors(10), trans = 'reverse') +
  theme(#legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust=1),
        panel.border = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "white"))
ggsave(path.out)

  #summarise(r = range(Rsquared))
  #round(2) %>%
  #kable()

>>>>

df <- df %>% 
  group_by(year) %>% 
  mutate(n.slag = lag.listw(knn.l, n)) %>% 
  group_by(BUA22CD) %>% 
  mutate(n.lag = dplyr::lag(n, n=1, order_by=year),
         n.l.slag = dplyr::lag(n.slag, n=1, order_by=year),
         n.London.lag = dplyr::lag(n.London, n=1, order_by = year),
         n.nearest.city.lag = dplyr::lag(n.nearest.city, n=1, order_by = year)) %>%
  ungroup()
```

@@

@@

# Lookup: BUA to regions

# source: https://www.data.gov.uk/dataset/aca8408d-e422-4ec2-a22f-9bd8beb1d549/built-up-area-to-region-december-2022-lookup-in-great-britain
path.lookup <- paste0(path, "/data/raw/Built_Up_Area_to_Region_(December_2022)_Lookup_in_Great_Britain.csv")

# 142 BUA in more than one regions
lookup.region <- read_csv(path.lookup)

df <- df %>% left_join(lookup.region %>% 
                         filter(WHOLE_PART=="Whole") %>% 
                         dplyr::select(BUA22CD, RGN22CD, RGN22NM), 
                       by = c("BUA22CD" = "BUA22CD")) %>% 
  #filter(is.na(RGN22CD))                                      #142 BUA in more than one regions dropped
  relocate(c(RGN22NM, RGN22CD), .after=BUA22NM) %>% 
  ungroup()

@@

# For growth:
df <- df %>% group_by(BUA22CD) %>% 
  mutate(#n.lag = dplyr::lag(n),
         growth = log(n/n.lag),
                    abs.growth = n - n.lag) %>% 
  relocate(n.lag, .after = n)

# out.path <- paste0(path, "/data/temp/df_for_bua_rf.csv")
# df <- read_csv(out.path)
# df %>% write_csv(file = out.path)

@@

# There are no lags for 1996, so it is removed
df <- df %>% filter(year > 1996) %>% 
  dplyr::select(BUA22CD, BUA22NM, RGN22NM, RGN22CD, year, n, n.l.slag, n.nearest.city.lag, 
                n.London.lag, City.of.Westminster, dist,) %>% 
  rename(London = City.of.Westminster)
```

@@

```{r}
#RNGversion("4.0.2")

set.seed(123)
indices <- CreateSpacetimeFolds(df, 
                                spacevar = "BUA22CD",
                                timevar = "year", 
                                k = 10,
                                seed = 123)

@@

start.time <- Sys.time()
model.all <- train(n ~ #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + n.nearest.city + year,
                     n.London.lag + n.nearest.city.lag + n.l.slag + year + London + dist,
                     #London.v + city.v + retail.v + n.l.slag + year,
                   data = df, 
                   trControl = tc,
                   method = "ranger",
                   na.action = na.omit,
                   #preProc = c("center", "scale"),
                   importance = "impurity")

stopCluster(cl)

end.time <- Sys.time()
time.taken <- round(end.time - start.time,2)
time.taken #7.3m

@@

# Random Forest 
# 
# 134528 samples
#      6 predictor
# 
# No pre-processing
# Resampling: Cross-Validated (10 fold) 
# Summary of sample sizes: 113520, 105938, 105938, 113505, 113505, 105938, ... 
# Resampling results across tuning parameters:
# 
#   mtry  splitrule   RMSE      Rsquared   MAE     
#   2     variance    81.30945  0.6421819  23.39696
#   2     extratrees  96.94266  0.5158416  26.31072
#   4     variance    78.74613  0.6597365  21.00403
#   4     extratrees  88.39690  0.5849950  25.41510
#   6     variance    79.57023  0.6514796  20.86435
#   6     extratrees  86.15949  0.6003589  25.47657
# 
# Tuning parameter 'min.node.size' was held constant at a value of 5
# RMSE was used to select the optimal model using the smallest value.
# The final values used for the model were mtry = 4, splitrule = variance and min.node.size = 5.

varimp_mars <- varImp(model.all) 
varimp_mars$importance <- varimp_mars$importance %>% 
  rownames_to_column() %>% 
  mutate(rowname = ifelse(rowname == "dist", "distance to the nearest city",
                          ifelse(rowname == "dist.retail", "distance to the nearest retail centre", 
                                 ifelse(rowname == "London", "distance to London",
                                        ifelse(rowname == "n.nearest.retail.lag", "Nearest retail centre's wensite density, t-1",
                                               ifelse(rowname == "n.London.lag", "London's wensite density, t-1",
                                                      ifelse(rowname == "n.l.slag", "spatial and temporal lag of wensite density",
                                                             ifelse(rowname == "n.nearest.city.lag", "Nearest city's wensite density, t-1", rowname)))))))) %>% 
  column_to_rownames()

path.out <- paste0(path, "/outputs/rf/figures/varimp_BUA.png")
ggplot(varimp_mars) + theme_minimal() + labs(x="") #+ theme(panel.background = element_rect(fill = "white"))
ggsave(path.out)
```

@@

```{r}
# Drop the 142 BUA in more than one regions
df <- df %>% filter(!is.na(RGN22CD))

#rm(list = setdiff(ls(), c("df", "indices")))

@@

# detectCores()
cl <- makePSOCKcluster(4)
registerDoParallel(cl)

@@

regions <- df %>% ungroup() %>% distinct(RGN22CD) 
regions <- as.vector(regions$RGN22CD)

@@

  set.seed(71)
  indices <- CreateSpacetimeFolds(df %>% filter(RGN22CD!=i), 
                                spacevar = "BUA22CD",
                                timevar = "year", 
                                k = k)

@@

  model.all <- train(n ~ n.London.lag + n.nearest.city.lag + n.l.slag + year + London + dist,
                   data = df %>% filter(RGN22CD!=i), 
                   trControl = tc,
                   method = "ranger",
                   na.action = na.omit,
                   #preProc = c("center", "scale"),
                   importance = "impurity")
  assign(paste0("model.all",as.character(i)), model.all) # object name = region not in the training data

  # file <- paste0("/hdd/tmp/regions/region_", as.character(i), ".RData")  
  # save(model.all, file = file)  # file name = region not in the training data
  rm(model.all)

@@

# create a list of the above model outputs
fit.model.all <- mget(ls(pattern = "^model.allE|^model.allN|^model.allW|^model.allS"))

# clear up
rm(list = ls(pattern = "^model.allE|^model.allN|^model.allW|^model.allS"))

@@

df.predictions <- data.frame()
for (i in regions){
  #for (j in regions[!regions == i]){
    pred <- predict(fit.model.all[names(fit.model.all)==paste0("model.all", i)], 
                    df[df$RGN22CD==i,])
    pred <- as.data.frame(pred)
    rownames(pred) <- c()
    assign(paste0("region.predict.model.", ".on.", i), pred)

    d <- paste0("region.predict.model.", ".on.", i)
    d <- cbind(get(d), df %>% filter(RGN22CD==i)) #%>% arrange(year, ladcd)

    d <- d %>% dplyr::select(RGN22CD, RGN22NM, BUA22CD, year, n, 1)

    colnames(d)[6] <- "predictions"
    d <- d %>% mutate(trained.on = i)
    df.predictions <- rbind(d, df.predictions)
  #}
}

df.predictions <- df.predictions %>% mutate(test.train = paste0(RGN22CD, ".", trained.on))

@@

# source: https://www.data.gov.uk/dataset/aca8408d-e422-4ec2-a22f-9bd8beb1d549/built-up-area-to-region-december-2022-lookup-in-great-britain
path.lookup <- paste0(path, "/data/raw/Built_Up_Area_to_Region_(December_2022)_Lookup_in_Great_Britain.csv")

# 142 BUA in more than one regions
lookup.region <- read_csv(path.lookup) %>% 
  dplyr::select(RGN22CD, RGN22NM) %>% 
  distinct()

path.out <- paste0(path, "/outputs/rf/test_regions_BUA.csv")

rf.year.all.metrics %>% 
  as.data.frame() %>%
  rownames_to_column(var = "metrics") %>% 
  pivot_longer(!metrics, names_to = "train.test") %>% 
  pivot_wider(names_from = metrics, values_from = value) %>% 
  separate(train.test, c("train", "test"), remove = F) %>% 
  left_join(lookup.region, by = c("train" = "RGN22CD")) %>% 
  rename(Region = RGN22NM) %>% 
  arrange(Rsquared) %>% 
  dplyr::select(Region, Rsquared) %>% 
  write_csv(path.out)
```
